\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-18 paper/conclusion.tex

\subsection{Future Work}
\todo{Jose: rewrite this}
\ProjectName{} opens up multiple avenues for further work. For instance, we
could extend our minimal/maximal imputation operators to consider global
information, such as the specific columns needed in operators higher up in the
query plan.  Our optimizer uses the same machine learning algorithm in all
instances of imputation operators. Integrating multiple possible algorithms
could allow for further fine-grained imputation, and honing the time complexity and
loss of these algorithms for an iterator model database could facilitate more
intelligent query plans and consistent interpretations of $\alpha$ across
strategies. For example, algorithms that learn in an online manner could
increase the efficiency of the system.  Furthermore, a multiple imputation
strategy could be followed for queries involving certain aggregates. Finally,
the underlying database for \ProjectName{} is far from a full-featured,
production database, so a natural step is to integrate imputation planning in a
widely-adopted production-quality database. This will likely expose further
opportunities for improvement.

>>>>

\subsection{Future Work}
\ProjectName{} opens up multiple avenues for further work. We sketch out some of these
directions.

\textbf{Cost model intepretation:} While the current system
provides a parameter ($\alpha$) to tradeoff between imputation quality and runtime,
there is no direct interpretation of its unit-less value. One natural extension to this work
is to reformulate the cost-model in an interpretable fashion.

\textbf{Imputation confidence:} Obtaining a confidence measure for query results
produced when missing values are imputed with a machine learning model is
another possible improvement to the system. In most cases, the data imputed
is unlikely to follow a standard distribution, which implies that any confidence
estimates will result from iterative approaches such as cross-validation\cite{kohavi1995study}.
In contrast to standard confidence calculations, there is the added complexity
of imputation taking place at various points in the plan and the composition of results
through standard query operations.

\textbf{Imputation operators:} We explored a subset of possible imputation operators.
In particular, we explored the two variants of the drop and impute operations. The system
could be extended to consider a broader family of operators. As the search space grows,
there will likely be additional steps needed in order to avoid sub-optimal plans.

\textbf{Adaptive query planning:} Currently, the imputation is local to a given query, meaning
no information is shared across queries. An intriguing direction would be to take advantage
of imputation and execution data generated by repeated queries. Note that this goes beyond
simply caching imputation results, and instead could entail operations such as extending
intermediate results with prior imputation values to improve accuracy, or pruning out
query plans which have shown to produce low-confidence imputations.

### 2017-02-18 paper/experiments.tex

\subsubsection{freeCodeCamp: 2016 New Coder Survey}
For our second set of experiments, we chose to use data collected
by freeCodeCamp as a part of a survey of new software developers
(both professional and amateur). freeCodeCamp is an open-source
community that intends to help users learn how to program. In addition,
to providing training material for users, it also pairs users with
nonprofits to develop software solutions. Their \textit{2016 New Coder Survey} consists of responses by over 15,000 people to 48 different
questions. The survey targeted users who were in one way or another
related to coding organizations.

>>>>

\subsubsection{freeCodeCamp: 2016 New Coder Survey}
For our second set of experiments, we chose to use data collected
by freeCodeCamp as a part of a survey of new software developers
(both professional and amateur)\cite{fcc-data}. freeCodeCamp is an open-source
community that intends to help users learn how to program. In addition,
to providing training material for users, it also pairs users with
nonprofits to develop software solutions. Their \textit{2016 New Coder Survey} consists of responses by over 15,000 people to 48 different
questions. The survey targeted users who were in one way or another
related to coding organizations.

### 2017-02-18 paper/introduction.tex
paper: minor word changes

Handling incorrect or dirty data is a complex and challenging problem for data scientists.
One way in which a dataset can be dirty is for parts of it to be missing altogether.
Missing data is one of the simpler variants of dirty data, but if handled naively, can still cause analyses to be incorrect.
To handle this problem, users may manually clean their dataset by performing
some statistical analysis to replace missing data elements with likely values.
This process is called \emph{imputation}.
Traditional imputation methods replace all missing values in a dataset to create a new clean dataset.
Although manual imputation solves the problem of missing data, in the age of big data it may be very expensive to run an imputation algorithm on an entire dataset.
Additionally, it may not even be necessary to completely clean the data to make it usable.
Some users may be willing to run queries on dirty data, simply ignoring the missing values, as long as they do not have to pay the cost of imputation.
Others may want to run queries on a subset of the data, so do not need to impute
every field in every record. Yet others may want to customize the 
imputation algorithm for the tradeoffs and demands of a particular domain.

>>>>

Handling incorrect or dirty data is a complex and challenging problem for analysts.
According to some estimates, data scientists spend between 50 and 80 percent of their
time performing tasks such as collecting and cleaning datasets\cite{data-science-cleaning}.
One way in which a dataset can be dirty is for parts of it to be missing altogether.
Missing data is one of the simpler variants of dirty data, but if handled naively, can still cause analyses to be incorrect.
To handle this problem, users may manually clean their dataset by performing
some statistical analysis to replace missing data elements with likely values.
This process is called \emph{imputation}.
Traditional imputation methods replace all missing values in a dataset to create a new clean dataset.
Although manual imputation solves the problem of missing data, in the age of big data it may be very expensive to run an imputation algorithm on an entire dataset.
Additionally, it may not even be necessary to completely clean the data to make it usable.
Some users may be willing to run queries on dirty data, simply ignoring the missing values, as long as they do not have to pay the cost of imputation.
Others may want to run queries on a subset of the data, and so do not need to impute
every field in every record. Yet others may want to customize the 
imputation algorithm for the tradeoffs and demands of a particular domain.
